SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-10-27 docs/questioning.tex
rewrite Figure 2 description

An attack starts with the malicious user sending an request either to
the server or to a web user.
In the first case 
consider a web application that accepts and redisplays data of
uncertain origin without appropriate validation and / or filtering.
An attacker could upload data containing a specially
crafted script (for example, in a blog comment)
to steal the cookies of the visiting users
or embed malicious {\sc sql} code to retrieve all
the password entries from a database.
In the second case, the attacker crafts a malicious script
into a {\sc url} and then tricks an end user to click on it
via social engineering (i.e. phishing). In this way
an unauthorized command is transmitted from a user to a
website where a user is considered as legitimate.

Then, the injected code reaches the vulnerable
application where it becomes a part of a value
represented by a program variable.
In the next step, the code becomes a part of a query that
finally reaches the database where it is executed.
Attacks like {\sc sql} injection and {\sc csrf}
end right there.

Before we proceed with the {\sc xss} attack path,
we should mention that there are two basic
{\sc xss} attack categories, namely:
{\it non-persistent} and {\it persistent}.
Non-persistent {\sc xss} attacks take place when the data provided
by a user, is used immediately by server-side scripts to parse
and display a page of results for and to that user, without
properly sanitizing the request. In the case of a
persistent {\sc xss} attack, malicious code
is saved by the server, and then permanently
displayed on the web site's pages that are displayed
to other users. Thus, in a non-persistent {\sc xss}
attack the injected code never reaches the database
and immediately becomes a part of the content that is
sent back to a web user. In the case of a persistent
{\sc xss}, the injected code that is stored on the server-side
re-enters the application's execution flow and afterwards it
becomes a part of the content that is actually sent back to
the user.

>>>>

A malicious user can initiate the attack through two different routes.
The attacker may use the browser of an innocent user as an attack
vehicle, through which the code will be injected in the application.
For example, the attacker could craft a malicious script into a {\sc url}
and then trick an end user to click on it via social engineering
(i.e., phishing). Alternatively, the attacker may be able to inject
directly the malicious code in the application. For example, consider
a web application that accepts and redisplays data of uncertain origin
without appropriate validation or filtering. An attacker could
upload data containing a specially crafted script (for example, in a
blog comment) to steal the cookies of the visiting users or embed
malicious {\sc sql} code to retrieve all the password entries from a
database.

Once the injected code reaches the vulnerable application, it becomes
a part of a value represented by a program variable. The target of the
attack determines the route from now on. In {\sc sql} injection and
{\sc csrf} attacks the injected code becomes part of a query that
finally reaches the database where it is executed.

{\sc xss} attacks fall ino two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request. Thus, in a non-persistent {\sc xss} attack the injected code
never reaches the database and immediately becomes a part of the
content that is sent back to a web user.

In persistent {\sc xss} attacks malicious code is saved by the server.
The injected code that is stored on the server-side re-enters the
application's execution flow and becomes part of the content that is
actually sent back afterwards to the user.
